---
title: "TFM data merged"
editor: visual
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = TRUE   
)
```

```{r}
library(readxl)
library(dplyr) 
library(purrr)
library(stringr)
library(janitor)
library(gdata)
library(tidyverse)
library(tidyr)
library(stringdist)
library(ggcorrplot)
library(ggplot2)
library(mice)
library(missForest)
library(plm)  
library(fastDummies)
library(fixest)
library(stringmagic)
library(plm)
library(lpirfs)
library(stargazer)
library(broom) 
library(kableExtra)
```

## Data input and Data cleaning from 2003 to 2023

```{r}
df2003 <- read_excel("SB2004_EU500.xlsx", skip = 5, na = "..") %>% clean_names() %>%
  mutate(year = 2003)
colnames(df2003) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_2003_million", "rd_growth_03_02_percent",
  "rd_growth_02_01_percent", "rd_growth_01_00_percent",

  "sales_2003_million", "sales_growth_03_02_percent",
  "sales_growth_02_01_percent", "sales_growth_01_00_percent",

  "employees_2003", "employees_growth_03_02_percent",

  "rd_intensity_2003_percent", "rd_intensity_2002_percent",

  "profitability_2003_percent",

  "rd_per_employee_2003_keuro", "rd_per_employee_2002_keuro",
  "rd_per_employee_cagr_3y_percent",

  "market_cap_2003_million", "market_cap_growth_03_02_percent",
  "market_cap_cagr_3y_percent",

  "year"
)
```

```{r}
df2004 <- read_excel("SB2005_EU700.xlsx", skip = 4, na = "..") %>%
  clean_names() %>%
  mutate(year = 2004)

colnames(df2004) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_investment_2004_million", 
  "rd_investment_change_04_03_percent",
  "rd_investment_change_03_02_percent",
  "rd_investment_change_02_01_percent",

  "net_sales_2004_million", 
  "net_sales_change_04_03_percent",
  "net_sales_change_03_02_percent",
  "net_sales_change_02_01_percent",

  "employees_2004", 
  "employees_change_04_03_percent",

  "rd_sales_ratio_2004_percent", 
  "rd_sales_ratio_2003_percent",

  "operating_profit_2004_percent_of_sales",

  "rd_per_employee_2004_k_euro", 
  "rd_per_employee_2003_k_euro",

  "market_cap_2004_million", 
  "market_cap_change_04_03_percent",

  "capex_2004_percent_of_sales",
  "capex_change_04_03_percent_of_sales",

  "year"
)
```

```{r}
df2005 <- read_excel("SB2006_EU1000.xlsx", skip = 4, na = " ") %>%
  clean_names() %>%
  mutate(year = 2005)

colnames(df2005) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_investment_2005_million", 
  "rd_investment_change_05_04_%", 
  "rd_investment_change_04_03_%", 
  "rd_investment_change_03_02_%",

  "net_sales_2005_million", 
  "net_sales_change_05_04_%", 
  "net_sales_change_04_03_%", 
  "net_sales_change_03_02_%",

  "employees_2005", 
  "employees_change_05_04_%", 

  "rd_sales_ratio_2005_%", 
  "rd_sales_ratio_2004_%",

  "operating_profit_2005_%_of_sales",

  "rd_per_employee_2005_ke", 
  "rd_per_employee_2004_ke",

  "market_cap_2005_million", 
  "market_cap_change_05_04_%",

  "capex_2005_%_of_sales", 
  "capex_2004_%_of_sales",
  
  "year"
)
```

```{r}
df2006 <- read_excel("SB2007_EU1000.xlsx", skip = 4, na = " ") %>%
  clean_names() %>%
  mutate(year = 2006)

colnames(df2006) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_investment_2006_million", "rd_investment_change_06_05_percent",
  "rd_investment_change_05_04_percent", "rd_investment_change_04_03_percent",

  "net_sales_2006_million", "net_sales_change_06_05_percent",
  "net_sales_change_05_04_percent", "net_sales_change_04_03_percent",

  "employees_2006", "employees_change_06_05_percent",

  "rd_sales_ratio_2006_percent", "rd_sales_ratio_2005_percent",

  "operating_profit_2006_percent_of_sales",

  "rd_per_employee_2006_k_euro", "rd_per_employee_2005_k_euro",

  "market_cap_2006_million", "market_cap_change_06_05_percent",

  "capex_2006_percent_of_sales", "capex_2005_percent_of_sales",

  "year"
)
```

```{r}
df2007 <- read_excel("SB2008_EU1000.xlsx", skip = 4, na = " ") %>%
  clean_names() %>%
  mutate(year = 2007)

colnames(df2007) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_investment_2007_million", 
  "rd_investment_change_07_06_percent", 
  "rd_investment_change_06_05_percent", 
  "rd_investment_change_05_04_percent",

  "net_sales_2007_million", 
  "net_sales_change_07_06_percent", 
  "net_sales_change_06_05_percent", 
  "net_sales_change_05_04_percent",

  "employees_2007", 
  "employees_change_07_06_percent", 

  "rd_sales_ratio_2007_percent", 
  "rd_sales_ratio_2006_percent",

  "operating_profit_2007_percent_of_sales",

  "rd_per_employee_2007_ke", 
  "rd_per_employee_2006_ke",

  "market_cap_2007_million","market_cap_change_07_06_percent_of_sales",
  "capex_2007_percent_of_sales",
  "capex_2006_percent_of_sales",

  "year"
)
```

```{r}
df2008 <- read_excel("SB2009_EU1000.xlsx", skip = 4, na = " ") %>%
  clean_names() %>%
  mutate(year = 2008)

colnames(df2008) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_investment_2008_million", "rd_investment_change_08_07_percent",
  "rd_investment_change_07_06_percent", "rd_investment_change_06_05_percent",

  "net_sales_2008_million", "net_sales_change_08_07_percent",
  "net_sales_change_07_06_percent", "net_sales_change_06_05_percent",

  "employees_2008", "employees_change_08_07_percent",

  "rd_sales_ratio_2008_percent", "rd_sales_ratio_2007_percent",

  "operating_profit_2008_percent_of_sales",

  "rd_per_employee_2008_ke", "rd_per_employee_2007_ke",

  "market_cap_2008_million","market_cap_change_08_07_percent", "capex_2008_percent_of_sales",
  "capex_2007_percent_of_sales",

  "year"
)
```

```{r}
df2009 <- read_excel("SB2010_EU1000.xlsx", skip = 4, na = " ") %>% clean_names() %>% mutate(year = 2009)
colnames(df2009) <- c(
  "rank", "company", "icb_sector", "country",
  
  "rd_investment_2009_million", "rd_investment_change_09_08_percent", 
  "rd_investment_change_08_07_percent", "rd_investment_change_07_06_percent",

  "net_sales_2009_million", "net_sales_change_09_08_percent", 
  "net_sales_change_08_07_percent", "net_sales_change_07_06_percent",

  "employees_2009", "employees_change_09_08_percent",

  "rd_sales_ratio_2009_percent", "rd_sales_ratio_2008_percent",

  "operating_profit_2009_percent_of_sales",

  "rd_per_employee_2009_k_euro", "rd_per_employee_2008_k_euro",

  "market_cap_2009_million", "market_cap_change_09_08_percent",

  "capex_2009_percent_of_sales", "capex_2008_percent_of_sales",

  "year"
)
```

```{r}
df2010 <- read_excel("SB2011_EU1000.xlsx", skip = 4, na = " ") %>% clean_names() %>% mutate(year = 2010)
colnames(df2010) <- c(
  "rank", "company", "icb_sector", "country",

  "rd_2010_million", "rd_growth_1y_percent",
  "rd_investment_change_09_08_percent", "rd_investment_change_08_07_percent",

  "sales_2010_million", "sales_growth_1y_percent",
  "sales_change_09_08_percent", "sales_change_08_07_percent",

  "employees_2010", "employees_growth_1y_percent",

  "rd_intensity_percent", "rd_sales_ratio_2009_percent",

  "profitability_percent",

  "rd_per_employee_2010_k_euro", "rd_per_employee_2009_k_euro",

  "market_cap_2010_million", "market_cap_growth_1y_percent",

  "capex_intensity_percent", "capex_2009_percent_of_sales",

  "year"
)
```

```{r}
df2011 <- read_excel("SB2012_EU1000.xlsx", skip = 2, na = " ") %>% clean_names() %>% mutate(year = 2011)
colnames(df2011) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2011_million", "rd_growth_1y_percent", "rd_cagr_3y_percent",

  "sales_2011_million", "sales_growth_1y_percent", "sales_cagr_3y_percent",

  "rd_intensity_percent",

  "capex_2011_million", "capex_growth_1y_percent", "capex_cagr_3y_percent",
  "capex_intensity_percent",

  "profit_2011_million", "profit_growth_1y_percent", "profit_cagr_3y_percent",
  "profitability_percent",

  "employees_2011", "employees_growth_1y_percent", "employees_cagr_3y_percent",

  "year"
)
```

```{r}
df2012 <- read_excel("SB2013_EU1000.xlsx", skip = 3, na = " ") %>% clean_names() %>% mutate(year = 2012)
colnames(df2012) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2012_million", "rd_growth_1y_percent", "rd_cagr_3y_percent",
  "rd_intensity_percent",

  "sales_2012_million", "sales_growth_1y_percent", "sales_cagr_3y_percent",

  "capex_2012_million", "capex_growth_1y_percent", "capex_cagr_3y_percent",
  "capex_intensity_percent",

  "profit_2012_million", "profit_growth_1y_percent", "profit_cagr_3y_percent",
  "profitability_percent",

  "employees_2012", "employees_growth_1y_percent", "employees_cagr_3y_percent",

  "year"
)
```

```{r}
df2013 <- read_excel("SB2014_EU1000.xlsx", skip = 1, na = " ") %>% clean_names() %>% mutate(year = 2013)
colnames(df2013) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2013_million", "rd_growth_1y_percent", "rd_cagr_3y_percent",
  "rd_intensity_percent",

  "sales_2013_million", "sales_growth_1y_percent", "sales_cagr_3y_percent",

  "capex_2013_million", "capex_growth_1y_percent", "capex_cagr_3y_percent",
  "capex_intensity_percent",

  "profit_2013_million", "profit_growth_1y_percent", "profit_cagr_3y_percent",
  "profitability_percent",

  "employees_2013", "employees_growth_1y_percent", "employees_cagr_3y_percent",

  "year"
)
```

```{r}
df2014 <- read_excel("SB2015_EU1000.xlsx", skip = 2, na = " ") %>% clean_names() %>% mutate(year = 2014)
colnames(df2014) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2014_million", "rd_growth_1y_percent", "rd_cagr_3y_percent",

  "sales_2014_million", "sales_growth_1y_percent", "sales_cagr_3y_percent",

  "rd_intensity_percent",

  "capex_2014_million", "capex_growth_1y_percent", "capex_cagr_3y_percent", "capex_intensity_percent",

  "profit_2014_million", "profit_growth_1y_percent", "profit_cagr_3y_percent",
  "profitability_percent",

  "employees_2014", "employees_growth_1y_percent", "employees_cagr_3y_percent",

  "year"
)
```

```{r}
df2015 <- read_excel("SB2016_EU1000.xlsx", skip = 2, na = " ") %>% clean_names() %>% mutate(year = 2015)
colnames(df2015) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2015_million", "rd_growth_1y_percent", "rd_cagr_3y_percent",

  "sales_2015_million", "sales_growth_1y_percent", "sales_cagr_3y_percent",

  "rd_intensity_percent",

  "capex_2015_million", "capex_growth_1y_percent", "capex_cagr_3y_percent", "capex_intensity_percent",

  "profit_2015_million", "profit_growth_1y_percent", "profit_cagr_3y_percent",
  "profitability_percent",

  "employees_2015", "employees_growth_1y_percent", "employees_cagr_3y_percent",

  "year"
)
```

```{r}
df2016 <- read_excel("SB2017_EU1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2016)
colnames(df2016) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2016_million", "rd_growth_1y_percent",

  "sales_2016_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2016_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2016_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2016_thousand", "employees_growth_1y_percent",

  "market_cap_2016_million", "market_cap_growth_1y_percent",

  "year"
)
df2016 <- df2016 %>%
  rename(employees_2016 = employees_2016_thousand) %>%  
  mutate(employees_2016 = employees_2016 * 1000)        
```

```{r}
df2017 <- read_excel("SB2018_EU1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2017)
colnames(df2017) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2017_million", "rd_growth_1y_percent",

  "sales_2017_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2017_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2017_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2017", "employees_growth_1y_percent",

  "market_cap_2017_million", "market_cap_growth_1y_percent",

  "year"
)
```

```{r}
df2018 <- read_excel("SB2019_EU1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2018)
colnames(df2018) <- c(
  "rank", "company", "country", "icb_sector",

  "rd_2018_million", "rd_growth_1y_percent",

  "sales_2018_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2018_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2018_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2018", "employees_growth_1y_percent",

  "market_cap_2018_million", "market_cap_growth_1y_percent",

  "year"
)
```

```{r}
df2019 <- read_excel("SB2020_EU+UK1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2019)
colnames(df2019) <- c(
  "rank", "company", "country", "region", "icb_sector",

  "rd_2019_million", "rd_growth_1y_percent",

  "sales_2019_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2019_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2019_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2019", "employees_growth_1y_percent",

  "market_cap_2019_million", "market_cap_growth_1y_percent",

  "year"
)
```

```{r}
df2020 <- read_excel("SB2021_EU1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2020)
colnames(df2020) <- c(
  "rank", "company", "country", "region", "icb_sector",

  "rd_2020_million", "rd_growth_1y_percent",

  "sales_2020_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2020_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2020_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2020", "employees_growth_1y_percent",

  "market_cap_2020_million", "market_cap_growth_1y_percent",

  "year"
)
```

```{r}
df2021 <- read_excel("SB2022_EU1000.xlsx",  na = " ") %>% clean_names() %>% mutate(year = 2021)
colnames(df2021) <- c(
  "rank", "company", "country", "region", "icb_sector",

  "rd_2021_million", "rd_growth_1y_percent",

  "sales_2021_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2021_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2021_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2021", "employees_growth_1y_percent",

  "market_cap_2021_million", "market_cap_growth_1y_percent",

  "year"
)
```

```{r}
df2022 <- read_excel("SB2023_EU1000.xlsx",  na = " ") %>% clean_names()
colnames(df2022) <- c(
  "rank", "company", "year", "country", "region", "icb_sector",

  "rd_2022_million", "rd_growth_1y_percent",

  "sales_2022_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2022_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2022_million", "profit_growth_1y_percent", "profitability_percent",

  "employees_2022", "employees_growth_1y_percent",

  "market_cap_2022_million", "market_cap_growth_1y_percent"
)
```

```{r}
df2023 <- read_excel("SB2024_EU800.xlsx",  na = " ") %>% clean_names() 
colnames(df2023) <- c(
  "rank", "company", "year", "country", "region", "icb_sector",

  "rd_2023_million", "rd_growth_1y_percent",

  "sales_2023_million", "sales_growth_1y_percent",

  "rd_intensity_percent",

  "capex_2023_million", "capex_growth_1y_percent", "capex_intensity_percent",

  "profit_2023_million", "profit_growth_1y_percent", "profitability_percent",

  "market_cap_2023_million", "market_cap_growth_1y_percent",

  "employees_2023", "employees_growth_1y_percent"
)
```

## Check the company

```{r}
list_of_dfs <- list( df2003,df2004,df2005, df2006, df2007, df2008, df2009,
                    df2010, df2011, df2012, df2013, df2014, df2015, df2016,
                    df2017, df2018, df2019, df2020, df2021, df2022, df2023)


list_of_dfs <- lapply(list_of_dfs, function(df) {
  df %>% mutate(company = toupper(company))
})

company_years <- lapply(list_of_dfs, function(df) {
  df %>% select(company, year) %>% distinct()
}) %>% bind_rows()

company_counts <- company_years %>%
  group_by(company) %>%
  summarise(n_years = n()) %>%
  arrange(desc(n_years))

consistent_companies <- company_counts %>%
  filter(n_years == 21)
```

To identify firms consistently present in the R&D rankings from 2003 to 2023, I merged annual datasets and standardized company names. Only 124 companies appeared in all 21 years. This small sample size limits the feasibility of in-depth statistical or econometric analysis, especially when further segmented by country, industry, or other variables.

## Company Persistence Analysis

```{r}
company_by_year <- lapply(list_of_dfs, function(df) {
  df %>% mutate(company = toupper(company)) %>% select(company, year) %>% distinct()
}) %>% bind_rows()

all_years <- sort(unique(company_by_year$year))

window_size <- 10

results <- list()

for (i in 1:(length(all_years) - window_size + 1)) {
  years_window <- all_years[i:(i + window_size - 1)]

  filtered <- company_by_year %>% filter(year %in% years_window)
  common_companies <- filtered %>%
    group_by(company) %>%
    summarise(n_years = n()) %>%
    filter(n_years == window_size)

  results[[i]] <- tibble(
    start_year = years_window[1],
    end_year = years_window[window_size],
    common_company_count = nrow(common_companies)
  )
}

window_stats <- bind_rows(results)
window_stats %>% filter(common_company_count >= 400) %>%
  arrange(desc(common_company_count))
```

To identify a robust analysis window, I applied a 10-year rolling window to the data from 2003 to 2023. For each window, I counted the number of companies consistently present in all 10 years. Among the options, the period 2010–2019 had the highest count with 417 consistent companies, providing a sufficiently large and stable sample. Therefore, I selected 2010–2019 as the target window for further analysis.

```{r}
target_years <- 2010:2019
company_years_target <- company_by_year %>%
  filter(year %in% target_years)
common_companies_10y <- company_years_target %>%
  group_by(company) %>%
  summarise(n_years = n()) %>%
  filter(n_years == length(target_years)) %>%
  pull(company)
```

To identify firms consistently present from 2010 to 2019, I filtered the data to include only those years and counted the number of appearances for each company. Companies with records in all 10 years were selected as the final sample for analysis.

## Merged data

```{r}
dfs <- list(df2010, df2011, df2012, df2013, df2014,
            df2015, df2016, df2017, df2018, df2019)

dfs <- lapply(dfs, function(df) {
  df %>%
    mutate(company = toupper(company)) %>%
    filter(company %in% common_companies_10y) |> 
    select(-year) 
})

dfs_named <- map2(dfs, target_years, function(df, y) {
  df %>%
    rename_with(~ paste0(., "_", y), .cols = -company)
})

wide_data <- reduce(dfs_named, ~ left_join(.x, .y, by = "company"))
```

```{r}
wide_data <- wide_data %>%
  mutate(across(where(~ !is.factor(.)), as.character)) %>%
  { meta <- select(., company, matches("^(rank|country|icb_sector|region|rd_growth_1y_percent|sales_growth_1y_percent|rd_intensity_percent|capex_intensity_percent|profitability_percent|employees_growth_1y_percent)_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = c("variable", "year"),
        names_pattern = "^(.*)_(\\d{4})$",
        values_to = "value"
      ) %>%
      pivot_wider(names_from = variable, values_from = value) %>%
      mutate(year = as.integer(year))

    rd <- select(., company, matches("^rd_\\d{4}_million_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^rd_\\d{4}_million_(\\d{4})$",
        values_to = "rd_million"
      ) %>%
      mutate(year = as.integer(year))
    
    sales <- select(., company, matches("^sales_\\d{4}_million_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^sales_\\d{4}_million_(\\d{4})$",
        values_to = "sales_million"
      ) %>%
      mutate(year = as.integer(year))

    employees <- select(., company, matches("^employees_\\d{4}_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^employees_\\d{4}_(\\d{4})$",
        values_to = "employees"
      ) %>%
      mutate(year = as.integer(year))
    
    capex <- select(., company, matches("^capex_\\d{4}_million_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^capex_\\d{4}_million_(\\d{4})$",
        values_to = "capex_million"
      ) %>%
      mutate(year = as.integer(year))

    capex_growth <- select(., company, matches("^capex_growth_1y_percent_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^capex_growth_1y_percent_(\\d{4})$",
        values_to = "capex_growth_1y_percent"
      ) %>%
      mutate(year = as.integer(year))

    profit <- select(., company, matches("^profit_\\d{4}_million_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^profit_\\d{4}_million_(\\d{4})$",
        values_to = "profit_million"
      ) %>%
      mutate(year = as.integer(year))

    profit_growth <- select(., company, matches("^profit_growth_1y_percent_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^profit_growth_1y_percent_(\\d{4})$",
        values_to = "profit_growth_1y_percent"
      ) %>%
      mutate(year = as.integer(year))

    market_cap <- select(., company, matches("^market_cap_\\d{4}_million_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^market_cap_\\d{4}_million_(\\d{4})$",
        values_to = "market_cap_million"
      ) %>%
      mutate(year = as.integer(year))

    market_cap_growth <- select(., company, matches("^market_cap_growth_1y_percent_\\d{4}$")) %>%
      pivot_longer(
        cols = -company,
        names_to = "year",
        names_pattern = "^market_cap_growth_1y_percent_(\\d{4})$",
        values_to = "market_cap_growth_1y_percent"
      ) %>%
      mutate(year = as.integer(year))

    others <- select(., -matches("^(rank|country|icb_sector|region|rd_\\d{4}_million|sales_\\d{4}_million|employees_\\d{4}|rd_growth_1y_percent|sales_growth_1y_percent|rd_intensity_percent|capex_intensity_percent|profitability_percent|employees_growth_1y_percent|capex_\\d{4}_million|capex_growth_1y_percent|profit_\\d{4}_million|profit_growth_1y_percent|market_cap_\\d{4}_million|market_cap_growth_1y_percent)_\\d{4}$"))

    long_data <- reduce(
      list(meta, rd, sales, employees, capex, capex_growth, profit, profit_growth, market_cap, market_cap_growth),
      ~ full_join(.x, .y, by = c("company", "year"))
    )

    left_join(long_data, others, by = "company")
  }

wide_data <- wide_data %>%
  select(
    -all_of(c(
      "rd_investment_change_09_08_percent_2010",
      "rd_investment_change_08_07_percent_2010",
      "sales_change_09_08_percent_2010",
      "sales_change_08_07_percent_2010",
      "rd_per_employee_2010_k_euro_2010",
      "rd_per_employee_2009_k_euro_2010",
      "rd_sales_ratio_2009_percent_2010",
      "capex_2009_percent_of_sales_2010" 
    )),
    -matches("^(rd|sales|capex|profit|employees)_cagr_3y_percent_\\d{4}$")
  )
```

```{r}
wide_data <- wide_data %>%
  mutate(region = ifelse(is.na(region), "EU", region)) %>%
  select(
    company, year, region, country, icb_sector, rank,
    rd_million, rd_growth_1y_percent, rd_intensity_percent,
    sales_million, sales_growth_1y_percent,
    employees, employees_growth_1y_percent,
    capex_million, capex_growth_1y_percent, capex_intensity_percent,
    profit_million, profit_growth_1y_percent, profitability_percent,
    market_cap_million, market_cap_growth_1y_percent
  )

wide_data <- wide_data %>%
  mutate(
    company = as.character(company),
    region = as.factor(region),
    country = as.factor(country),
    icb_sector = as.factor(icb_sector),

    across(
      .cols = c(
        rank,
        rd_million, rd_growth_1y_percent, rd_intensity_percent,
        sales_million, sales_growth_1y_percent,
        employees, employees_growth_1y_percent,
        capex_million, capex_growth_1y_percent, capex_intensity_percent,
        profit_million, profit_growth_1y_percent, profitability_percent,
        market_cap_million, market_cap_growth_1y_percent
      ),
      .fns = ~ as.numeric(.)
    ),
    year = as.integer(year)
  )
wide_data <- wide_data %>%
  mutate(
    icb_sector = gsub("\\s*\\([^\\)]+\\)", "", icb_sector), 
    icb_sector = str_trim(icb_sector) 
  )
wide_data <- wide_data %>%
  mutate(icb_sector = icb_sector %>%
           str_to_title() %>%
           str_trim() %>%
           str_squish())

wide_data <- wide_data %>%
  mutate(icb_sector = case_when(
    icb_sector == "Aerospace & Defence" ~ "Aerospace & Defense",
    icb_sector == "Gas, Water & Multiutilities" ~ "Gas, Water & Multi-Utilities",
    TRUE ~ icb_sector
  ))
wide_data$country <- as.character(wide_data$country)  
wide_data$country <- ifelse(wide_data$country == "The Netherlands", 
                            "Netherlands", 
                            wide_data$country)
```

I constructed a balanced panel dataset for the period 2010–2019 by selecting companies that appeared in all ten years. I merged company-level data across years and reshaped key variables (e.g., R&D, sales, employees) into long format. I removed redundant or irrelevant columns, set missing region values to "EU", and reordered the variables to facilitate further analysis.

## Descriptive analysia

1.Number of companies by country

```{r}
company_count_country <- wide_data %>%
  group_by(country) %>%
  summarise(n_companies = n_distinct(company)) %>%
  arrange(desc(n_companies))

ggplot(company_count_country, aes(x = reorder(country, n_companies), y = n_companies)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n_companies), hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(
    title = "Number of Companies by Country",
    x = "Country",
    y = "Number of Companies"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +  
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  
```

The bar chart shows the number of companies per country that consistently appeared in the EU R&D ranking from 2010 to 2019. Germany leads with 107 firms, followed by the UK (77), France (56), and Sweden (34). These top contributors reflect the strong and stable innovation presence of these countries in the EU during this decade. Other countries like Italy, Finland, and Denmark also maintain a noticeable presence, while the remaining countries contribute relatively fewer consistent players.

2.Number of companies by industry (ICB sector)

```{r}
wide_data$icb_sector <- dplyr::recode(wide_data$icb_sector,
  "Telecommunications Equipment" = "Technology Hardware & Equipment",
  "Computer Hardware"              = "Technology Hardware & Equipment",
  "Semiconductors"                 = "Technology Hardware & Equipment",

  "Electrical Components & Equipment" = "Electronic & Electrical Equipment",
  "Electronic Equipment"              = "Electronic & Electrical Equipment",

  "Pharmaceuticals" = "Pharmaceuticals & Biotechnology",
  "Biotechnology"   = "Pharmaceuticals & Biotechnology",

  "Commercial Vehicles & Trucks" = "Industrial Engineering",
  "Industrial Machinery"         = "Industrial Engineering",

  "Software"         = "Software & Computer Services",
  "Computer Services" = "Software & Computer Services",
  "Other Financials" = "Financial Services"
)


company_count_industry <- wide_data %>%
  group_by(icb_sector) %>%
  summarise(n_companies = n_distinct(company)) %>%
  arrange(desc(n_companies))

ggplot(company_count_industry, aes(x = reorder(icb_sector, n_companies), y = n_companies)) +
  geom_segment(aes(xend = icb_sector, y = 0, yend = n_companies), color = "gray70") +
  geom_point(color = "darkgreen", size = 3) +
  geom_text(aes(label = n_companies), hjust = -0.8, size = 3.2) + 
  coord_flip() +
  labs(
    title = "Number of Companies by Industry (ICB Sector)",
    x = "ICB Sector",
    y = "Number of Companies"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.y = element_text(size = 7)
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)))  

```

The dot plot illustrates the number of companies in each ICB sector that consistently appeared in the EU R&D rankings from 2010 to 2019. The leading sectors are Industrial Engineering (57 firms), Industrial Machinery (47), and Pharmaceuticals & Biotechnology (44), highlighting the central role of these sectors in sustained innovation.

3.Average number of employees and net sales over time

```{r}
avg_stats <- wide_data %>%
  group_by(year) %>%
  summarise(
    avg_employees = mean(employees, na.rm = TRUE),
    avg_sales = mean(sales_million, na.rm = TRUE)
  )

ggplot(avg_stats, aes(x = year, y = avg_employees)) +
  geom_line(color = "blue", size = 1.2) +
  labs(title = "Average Number of Employees Over Time", x = "Year", y = "Employees") +
  scale_x_continuous(breaks = avg_stats$year) +  
  theme_minimal()

ggplot(avg_stats, aes(x = year, y = avg_sales)) +
  geom_line(color = "purple", size = 1.2) +
  labs(title = "Average Net Sales Over Time", x = "Year", y = "Net Sales (Million €)") +
  scale_x_continuous(breaks = avg_stats$year) + 
  theme_minimal()
```

Employees: The average number of employees remained relatively stable until around 2015, with minor fluctuations. From 2016 onwards, there was a noticeable increase in employment, indicating that firms were scaling up operations. This may reflect not just recovery but also structural growth in the industries covered.

Net Sales: The average net sales showed steady growth from 2010 to 2012, followed by fluctuations between 2013 and 2016. After 2016, net sales increased consistently and significantly, peaking in 2019. This trend suggests a post-crisis recovery and a likely expansion period in the EU corporate landscape during the late 2010s.

4.Analyze Rank Stability Across Companies (2010–2019):

```{r}
wide_data <- wide_data %>%
  mutate(rank = as.numeric(rank))
rank_stability <- wide_data %>%
  group_by(company) %>%
  summarise(
    avg_rank = mean(rank, na.rm = TRUE),  
    sd_rank = sd(rank, na.rm = TRUE)     
  ) %>%
  arrange(sd_rank)  

head(rank_stability, 10)

top_stable <- rank_stability %>% slice_min(sd_rank, n = 20)

ggplot(top_stable, aes(x = reorder(company, sd_rank), y = sd_rank)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Top 20 Most Stable Companies by Rank (2010–2019)",
    x = "Company",
    y = "Standard Deviation of Rank"
  ) +
  theme_minimal()
```

As shown in the chart, Volkswagen stands out with the most stable position, consistently ranked first across the entire period (SD = 0). Other firms like Daimler, Siemens, and Robert Bosch also show strong rank stability, indicating sustained investment and leadership in R&D.

On the other hand, companies such as Nokia, Evonik Industries, and L’Oréal demonstrate relatively higher rank variability, reflecting either fluctuations in R&D spending or changes in their relative position in the European innovation landscape.

This analysis highlights a core group of firms with long-term commitment to R&D, which can be important for further modeling or policy discussion regarding consistent innovation performance.

```{r}
cor_data <- wide_data %>%
  select(
    rd_million, rd_growth_1y_percent, rd_intensity_percent,
    sales_million, sales_growth_1y_percent,
    employees, employees_growth_1y_percent,
    capex_million, capex_growth_1y_percent, capex_intensity_percent,
    profit_million, profit_growth_1y_percent, profitability_percent,
    market_cap_million, market_cap_growth_1y_percent,rank
  ) %>%
  drop_na()  
cor_matrix <- cor(cor_data, use = "complete.obs")
ggcorrplot(
  cor_matrix,
  method = "circle",
  type = "lower",
  lab = TRUE,
  lab_size = 3,
  colors = c("red", "white", "blue"),
  title = "Correlation Matrix of Key Firm-Level Variables",
  ggtheme = theme_minimal()
)
```

R&D and Net Sales: There is a moderate positive correlation between R&D expenditure (rd_million) and net sales (sales_million) at 0.58, suggesting that firms with higher R&D investment tend to achieve higher sales revenue. Similarly, R&D is positively correlated with the number of employees (employees, 0.61), indicating that R&D-intensive firms typically have larger workforces.

Profit and Market Capitalization: The correlation between profit (profit_million) and market capitalization (market_cap_million) is 0.75, showing that more profitable firms generally have higher market value.

Capital Expenditure and Sales: Capital expenditure (capex_million) shows a strong positive correlation with net sales (sales_million) at 0.88, suggesting that firms with higher sales volumes are also more likely to make substantial capital investments.

Negative Correlations with Rank: The variable rank (where a lower number indicates a better ranking) is negatively correlated with key performance indicators like R&D, sales, and employees. This aligns with expectations—top-ranked firms tend to have more resources, higher output, and greater investment in R&D.

## Input datasets
```{r}
patents_data <- read.csv("patents count.csv")
long_patents_data <- patents_data %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "year",
    names_prefix = "X",
    values_to = "patent_count"
  ) %>%
  mutate(year = as.integer(year))  
merged_data <- left_join(wide_data, long_patents_data, by = c("company", "year"))

```

## Missing value analysis
```{r}
missing_percent <- round(colMeans(is.na(merged_data)) * 100, 2)
missing_percent
```

Due to the fact that the variable market_cap_growth_1y_percent (one-year market capitalization growth rate) has a high proportion of missing values (57.34%) and is a derived metric based on market_cap_million (market capitalization in million euros), it was excluded from further analysis. Including a variable with such a large amount of missing data could compromise the robustness of the analysis, and since the growth rate can be calculated from existing market cap values, retaining it was deemed unnecessary. Removing this variable helps simplify the dataset and maintain analytical clarity.

```{r}
merged_data <- merged_data |> 
  select(-market_cap_growth_1y_percent,-market_cap_million)
```

Given the presence of missing values across several key financial variables—such as capex_growth_1y_percent (23.12%), capex_million (19.59%), profit_million (11.22%), and profit_growth_1y_percent (12.11%)—I decided to perform imputation rather than removing the affected observations or variables. Since many of these indicators are important for the analysis, imputing missing values helps preserve the dataset’s completeness and analytical value. 

To address the 19.59% missing rate in capex_million, imputation was applied in order to preserve analytical integrity and avoid unnecessary data loss:
```{r}
ggplot(merged_data, aes(log10(capex_million + 1))) +
  geom_histogram(binwidth = 0.1, color = "skyblue3", fill = "skyblue") +
  ggtitle("Log-transformed Capex Distribution") +
  xlab("Log10(Capex + 1)") +
  ylab("Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 18))
```
Among several variables with missing values, capex_million (capital expenditure) was selected as a representative example for imputation comparison due to its relatively high missing rate (19.59%) and its importance as a financial indicator of corporate innovation investment. By focusing on this variable, I aim to evaluate and compare the effectiveness of different imputation strategies and their implications for subsequent data modeling and analysis.

## Selection of imputation strategy
```{r}
value_imputation <- merged_data |>
  mutate(
    original = capex_million,
    imputed_zero = if_else(is.na(capex_million), 0, capex_million),
    imputed_mean = if_else(is.na(capex_million), mean(capex_million, na.rm = TRUE), capex_million),
    imputed_median = if_else(is.na(capex_million), median(capex_million, na.rm = TRUE), capex_million)
  )

variables <- c("original", "imputed_zero", "imputed_mean", "imputed_median")
titles <- c("Original Capex Distribution", "Zero Imputation", "Mean Imputation", "Median Imputation")
colors <- c("skyblue3", "#15ad4f", "#6a6ad9", "#e65100")

value_imputation_long <- value_imputation %>%
  pivot_longer(all_of(variables), names_to = "method", values_to = "value")

plots <- value_imputation_long %>%
  mutate(title = factor(method, levels = variables, labels = titles)) %>%
  ggplot(aes(x = log10(value + 1), fill = title)) +
  geom_histogram(binwidth = 0.1, color = "#808080") +
  facet_wrap(~title, scales = "free_y") +
  scale_fill_manual(values = colors) +
  labs(x = "Log10(Capex + 1)", y = "Count") +
  theme_classic() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

print(plots)

```
These imputation methods differ significantly from the original distribution of capex_million. Zero imputation artificially creates a large number of zero values, severely distorting the distribution. Mean and median imputations, while simple and commonly used, concentrate the imputed values at a single point, resulting in unnatural peaks and a loss of the original distribution shape. Therefore, these methods are not suitable for highly skewed financial data.

Mice imputation:
```{r}
merged_numeric <- merged_data %>%
  select(capex_million, rd_million, sales_million, employees, profit_million)

mice_imputed <- data.frame(
  original = merged_data$capex_million,
  imputed_pmm = complete(mice(merged_numeric, m = 5, method = "pmm", seed = 123))$capex_million,
  imputed_cart = complete(mice(merged_numeric, m = 5, method = "cart", seed = 123))$capex_million,
  imputed_lasso = complete(mice(merged_numeric, m = 5, method = "lasso.norm", seed = 123))$capex_million
)

mice_imputed$imputed_rf <- complete(mice(merged_numeric, m = 5, method = "rf", seed = 123))$capex_million

variables <- c("original", "imputed_pmm", "imputed_cart", "imputed_lasso", "imputed_rf")
titles <- c(
  "Original Capex Distribution",
  "PMM-imputed Distribution",
  "CART-imputed Distribution",
  "Lasso-imputed Distribution",
  "RF-imputed Distribution"
)
colors_fill <- c("skyblue", "#15ad4f", "#6a6ad9", "#e65100", "#ffc107")

mice_imputed_long <- mice_imputed %>%
  pivot_longer(all_of(variables), names_to = "method", values_to = "value")

plots <- mice_imputed_long %>%
  mutate(title = factor(method, levels = variables, labels = titles)) %>%
  ggplot(aes(x = log10(value + 1), fill = title)) +
  geom_histogram(binwidth = 0.1, color = "#808080", position = "identity") +
  facet_wrap(~title, scales = "free_y") +
  scale_fill_manual(values = colors_fill) +
  labs(x = "Log10(Capex + 1)", y = "Count") +
  theme_classic() +
  theme(legend.position = "none", plot.title = element_text(size = 18))

print(plots)

```

Miss Random Forest :
```{r}
missForest_result <- missForest(as.data.frame(merged_numeric), ntree = 15)

missForest_imputed <- data.frame(
  original = merged_numeric$capex_million,
  imputed_missForest = missForest_result$ximp$capex_million
)

variables <- c("original", "imputed_missForest")
titles <- c("Original Capex Distribution", "MissForest-imputed Distribution")
colors_fill <- c("skyblue", "#FFD700")
colors_border <- c("skyblue3", "#808080")

missForest_imputed_long <- missForest_imputed %>%
  pivot_longer(all_of(variables), names_to = "method", values_to = "value")

plots <- missForest_imputed_long %>%
  mutate(title = factor(method, levels = variables, labels = titles)) %>%
  ggplot(aes(x = log10(value + 1), fill = title)) +
  geom_histogram(binwidth = 0.1, color = "#808080", position = "identity") +
  facet_wrap(~title, scales = "free_y") +
  scale_fill_manual(values = colors_fill) +
  theme_classic() +
  theme(legend.position = "none")

print(plots)
```

We applied and compared several imputation methods to handle the missing values in the capex_million variable, including Zero imputation, Mean imputation, Median imputation, Predictive Mean Matching (PMM), CART (Classification and Regression Trees), Lasso regression, Random Forest (RF), and MissForest.

The simpler methods—Zero, Mean, and Median imputation—introduced noticeable distortion. Zero imputation artificially inflated the frequency of zero values, while mean and median imputation caused unrealistic peaks at single values, which misrepresents the true underlying distribution.

Advanced methods such as PMM, CART, and Lasso performed better, but still introduced some smoothing or loss of variance compared to the original data. While these approaches attempt to model relationships between variables, they may not fully capture the complex, non-linear structure of financial data.

Among all, Random Forest (RF) imputation stood out as the most effective. It maintained the skewness, variability, and distribution shape of the original capex_million variable most accurately. This is likely due to RF's ensemble learning approach, which can model interactions and non-linearities effectively without overfitting.

Therefore, RF was selected as the most appropriate imputation method for this analysis, especially considering the financial nature and skewed distribution of the variable.

## Data imputation
```{r}
vars_to_impute <- c(
  "rd_growth_1y_percent", "rd_intensity_percent", "sales_million", 
  "sales_growth_1y_percent", "employees", "employees_growth_1y_percent", 
  "capex_million", "capex_growth_1y_percent", "capex_intensity_percent", 
  "profit_million", "profit_growth_1y_percent", "profitability_percent"
)

merged_numeric <- merged_data %>% select(all_of(vars_to_impute))

imputed_rf <- mice(merged_numeric, m = 5, method = "rf", seed = 123)

imputed_data_rf <- complete(imputed_rf, 1)

merged_data[vars_to_impute] <- imputed_data_rf

```

```{r}
round(colMeans(is.na(merged_data)) * 100, 2)
```
After imputation, we confirmed that all target variables no longer contain missing values, ensuring the dataset is complete and ready for subsequent analysis.

```{r}
merged_data <- merged_data %>% 
  arrange(company, year) %>% 
  mutate(
    patent_intensity_sales = patent_count / sales_million
  )
```

## Cross sectional analysis

### 1.by industry

H₀: Top20% ≤ Others
H₁: Top20% > Others
```{r}
data_2019 <- merged_data %>% filter(year == 2019)
data_2019 <- data_2019 %>%
  mutate(log_rd = log1p(rd_intensity_percent))
industry_avg_log <- data_2019 %>%
  group_by(icb_sector) %>%
  summarise(mean_log_rd = mean(log_rd, na.rm = TRUE))

threshold <- quantile(industry_avg_log$mean_log_rd, 0.80, na.rm = TRUE)

top20_industries <- industry_avg_log %>%
  filter(mean_log_rd >= threshold) %>%
  pull(icb_sector)
data_2019 <- data_2019 %>%
  mutate(industry_group = ifelse(icb_sector %in% top20_industries, "Top20%", "Others"))
data_2019$industry_group <- factor(data_2019$industry_group, levels = c("Top20%", "Others"))

t.test(patent_intensity_sales ~ industry_group, data = data_2019, alternative = "greater")

wilcox.test(sales_growth_1y_percent ~ industry_group, data = data_2019, alternative = "greater")


```

```{r}
 desc_stats <- data_2019 %>% 
  group_by(industry_group) %>% 
  summarise(
    n_firms           = n(),
    mean_pat_int      = mean(patent_intensity_sales,  na.rm = TRUE),
    sd_pat_int        = sd(patent_intensity_sales,    na.rm = TRUE),
    mean_sales_growth = mean(sales_growth_1y_percent, na.rm = TRUE),
    sd_sales_growth   = sd(sales_growth_1y_percent,   na.rm = TRUE),
    .groups = "drop"
  )

pat_test   <- t.test(patent_intensity_sales ~ industry_group,
                     data = data_2019, alternative = "greater") %>% tidy()

sales_test <- wilcox.test(sales_growth_1y_percent ~ industry_group,
                          data = data_2019, alternative = "greater") %>% tidy()

test_rows <- tibble(
  industry_group     = c("t-test (patent intensity)", "Wilcoxon (sales growth)"),
  n_firms            = NA,                
  mean_pat_int       = c(pat_test$estimate, sales_test$estimate),
  sd_pat_int         = c(pat_test$statistic, sales_test$statistic),
  mean_sales_growth  = c(pat_test$p.value,  sales_test$p.value),
  sd_sales_growth    = NA
)

results_table <- bind_rows(desc_stats, test_rows)

stargazer(results_table,
          type   = "text",     
          title  = "Top-20 % Industries vs. Others (2019) ",
          summary = FALSE,
          digits = 2,
          align  = TRUE,
          covariate.labels = c(
            "Group / Test", "N firms",
            "Mean Patent Intensity", "SD Patent Intensity",
            "Mean Sales Growth (%)", "SD Sales Growth (%)"
          ))

```

Based on 2019 data, firms in the top 20% of R&D-intensive industries show significantly stronger innovation output and growth performance compared to those in other industries. A Welch two-sample t-test indicates that their average patent intensity (patents per million in sales) is significantly higher (p = 0.005), with a mean of 0.13 versus just 0.02 in other industries—nearly an eightfold difference. Similarly, the Wilcoxon rank-sum test shows that sales growth is also significantly higher in the top 20% group (p < 0.001), with an average growth rate of 265%, compared to 5% in the rest. However, the large standard deviation in sales growth (around 3300%) suggests substantial variation, likely influenced by a few extreme high-growth firms. Overall, these findings support the hypothesis that firms in more R&D-intensive industries tend to be more innovative and achieve better growth outcomes.

### 2.by country

H₀: Top20% ≤ Others
H₁: Top20% > Others
```{r}
country_avg_log <- data_2019 %>%
  group_by(country) %>%
  summarise(mean_log_rd = mean(log_rd, na.rm = TRUE))

threshold <- quantile(country_avg_log$mean_log_rd, 0.80, na.rm = TRUE)

top20_countries <- country_avg_log %>%
  filter(mean_log_rd >= threshold) %>%
  pull(country)

data_2019 <- data_2019 %>%
  mutate(country_group = ifelse(country %in% top20_countries, "Top20%", "Others"))

data_2019$country_group <- factor(data_2019$country_group, levels = c("Top20%", "Others"))
t.test(patent_intensity_sales ~ country_group, data = data_2019, alternative = "greater")

wilcox.test(sales_growth_1y_percent ~ country_group, data = data_2019, alternative = "greater")
```

```{r}
desc_stats <- data_2019 %>% 
  group_by(country_group) %>% 
  summarise(
    n_firms           = n(),
    mean_pat_int      = mean(patent_intensity_sales,  na.rm = TRUE),
    sd_pat_int        = sd(patent_intensity_sales,    na.rm = TRUE),
    mean_sales_growth = mean(sales_growth_1y_percent, na.rm = TRUE),
    sd_sales_growth   = sd(sales_growth_1y_percent,   na.rm = TRUE),
    .groups = "drop"
  )

pat_test   <- t.test(patent_intensity_sales ~ country_group,
                     data = data_2019, alternative = "greater") %>% tidy()

sales_test <- wilcox.test(sales_growth_1y_percent ~ country_group,
                          data = data_2019, alternative = "greater") %>% tidy()

test_rows <- tibble(
  country_group      = c("t-test (patent intensity)", "Wilcoxon (sales growth)"),
  n_firms            = NA,
  mean_pat_int       = c(pat_test$estimate, sales_test$estimate),
  sd_pat_int         = c(pat_test$statistic, sales_test$statistic),
  mean_sales_growth  = c(pat_test$p.value,   sales_test$p.value),
  sd_sales_growth    = NA
)
results_table <- bind_rows(desc_stats, test_rows)

stargazer(results_table,
          type   = "text",     
          summary = FALSE,
          title  = "Top-20 % Countries vs. Others (2019) ",
          digits = 2,
          align  = TRUE,
          covariate.labels = c(
            "Group / Test", "N firms",
            "Mean Patent Intensity", "SD Patent Intensity",
            "Mean Sales Growth (%)", "SD Sales Growth (%)"
          ))
```

Firms in the top 20% of R&D-intensive countries did not show significantly higher patent intensity or sales growth compared to others. The t-test (p = 0.061) and Wilcoxon test (p = 0.125) both suggest no statistically significant differences, implying that the impact of country-level R&D intensity on firm outcomes may be weaker or less direct.

These empirical results highlight a notable contrast between industry- and country-level effects of R&D intensity. While firms in top R&D-intensive industries exhibit significantly higher innovation output and sales growth, no such relationship is observed at the country level. This suggests that the transmission of R&D to firm performance may be more strongly shaped by industry-specific technological dynamics than by broader national R&D environments.

## OLS 

1.R&D intensity
```{r}
panel_data <- merged_data %>%
  arrange(company, year) %>%
  group_by(company) %>%
  mutate(
    size_lag1  = dplyr::lag(sales_million, order_by = year)
  ) %>%
  mutate(
    year_lag1   = dplyr::lag(year, order_by = year)
  ) %>%
  mutate(
    rd_lag1     = dplyr::lag(rd_intensity_percent, order_by = year)
  ) %>%
  ungroup() 
panel_data<-panel_data |> 
   mutate(
    growth = log(sales_million) - log(size_lag1)
  ) 
panel_data <- panel_data %>%
  group_by(company) %>%
  mutate(growth_lag1 = dplyr::lag(growth, order_by = year)) %>%
  ungroup()

ols_model <- feols(
  growth ~ rd_lag1 + growth_lag1 + size_lag1 | icb_sector + country + year,
  data = panel_data
)
summary(ols_model)
```

The OLS regression results show that lagged R&D intensity has a positive and statistically significant effect on firm-level sales growth (coefficient = 0.0000215, p < 0.001). This means that a 1 percentage point increase in R&D intensity leads to an approximate 0.00215% increase in sales growth. Although the magnitude is small, the effect is meaningful in a log-linear model and may accumulate over time.

In contrast, lagged firm growth is significantly negative (–0.55), indicating a mean-reversion effect—firms that grew faster in the previous period tend to slow down subsequently. Lagged firm size also shows a weakly negative effect on growth, though it is not statistically significant at the 5% level.

The model includes industry, country, and year fixed effects, and explains approximately 42% of the variation in firm growth, with standard errors clustered at the industry level.

2.R&D growth
```{r}
panel_data1 <- merged_data %>%
  arrange(company, year) %>%
  group_by(company) %>%
  mutate(
    size_lag1   = dplyr::lag(sales_million, order_by = year),
    year_lag1   = dplyr::lag(year, order_by = year),
    rd_lag1     = dplyr::lag(rd_growth_1y_percent, order_by = year) 
  ) %>%
  ungroup() %>%
  mutate(
    growth = log(sales_million) - log(size_lag1)
  ) %>%
  group_by(company) %>%
  mutate(
    growth_lag1 = dplyr::lag(growth, order_by = year)
  ) %>%
  ungroup()

ols_model_growth <- feols(
  growth ~ rd_lag1 + growth_lag1 + size_lag1 | icb_sector + country + year,
  data = panel_data1
)
summary(ols_model_growth)
```

3.R&D log level
```{r}
panel_data2 <- merged_data %>%
  arrange(company, year) %>%
  group_by(company) %>%
  mutate(
    size_lag1   = dplyr::lag(sales_million, order_by = year),
    year_lag1   = dplyr::lag(year, order_by = year),
    rd_log      = log1p(rd_million),  
    rd_lag1     = dplyr::lag(rd_log, order_by = year)  
  ) %>%
  ungroup() %>%
  mutate(
    growth = log(sales_million) - log(size_lag1)
  ) %>%
  group_by(company) %>%
  mutate(
    growth_lag1 = dplyr::lag(growth, order_by = year)
  ) %>%
  ungroup()

ols_model_log <- feols(
  growth ~ rd_lag1 + growth_lag1 + size_lag1 | icb_sector + country + year,
  data = panel_data2
)
summary(ols_model_log)
```

## OLS-IV

1.R&D intensity
```{r}
panel_data <- panel_data %>%
  group_by(company) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(
    patent_lag1 = dplyr::lag(patent_intensity_sales, order_by = year)
  ) %>%
  ungroup()

iv_model <- feols(
  growth ~ 1 + growth_lag1 + size_lag1 | icb_sector + country + year | rd_lag1 ~ patent_lag1,
  data = panel_data
)
summary(iv_model)
```

The 2SLS regression addresses potential endogeneity in lagged R&D intensity (rd_lag1) by using lagged patent intensity (patent_lag1) as an instrument. The first-stage F-statistic is very high (30,168.7, p < 0.001), confirming that the instrument is highly relevant. The Wu-Hausman test rejects the null of exogeneity (p < 0.001), justifying the need for instrumental variable estimation.

In the second stage, the coefficient on fitted R&D intensity (fit_rd_lag1) is 0.0000266, which remains positive and marginally significant (p = 0.0608). This suggests that a 1 percentage point increase in R&D intensity leads to an estimated 0.00266% increase in firm sales growth. 

Lagged firm growth again shows a strong negative effect (–0.55), reinforcing the earlier finding of mean reversion, and firm size remains weakly negative and not statistically significant at conventional levels.

Overall, the IV results provide additional support for a positive, though modest, causal effect of R&D intensity on firm-level sales growth, and strengthen confidence in the interpretation of R&D as a productive investment. 

2.R&D growth
```{r}
panel_data1 <- panel_data1 %>%
  group_by(company) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(
    patent_lag1 = dplyr::lag(patent_intensity_sales, order_by = year)
  ) %>%
  ungroup()

iv_model_growth <- feols(
  growth ~ 1 + growth_lag1 + size_lag1 | icb_sector + country + year | rd_lag1 ~ patent_lag1,
  data = panel_data1
)

summary(iv_model_growth)

```

3.R&D log level
```{r}
panel_data2 <- panel_data2 %>%
  mutate(rd_log = log1p(rd_million)) %>%
  group_by(company) %>%
  arrange(year, .by_group = TRUE) %>%
  mutate(
    patent_lag1 = dplyr::lag(patent_intensity_sales, order_by = year)
  ) %>%
  ungroup()

iv_model_log <- feols(
  growth ~ 1 + growth_lag1 + size_lag1 | icb_sector + country + year | rd_lag1 ~ patent_lag1,
  data = panel_data2
)
summary(iv_model_log)
```

## GMM

1.R&D intensity
```{r}
library(plm)
merged_data <- merged_data %>%
  mutate(
    company = as.factor(company),
    year = as.integer(year)
  )

pdata <- pdata.frame(merged_data, index = c("company", "year"))
pdata <- na.omit(pdata)
gmm_model <- plm::pgmm(
  sales_growth_1y_percent ~ plm::lag(sales_growth_1y_percent, 1) + plm::lag(sales_million, 0:1) + plm::lag(rd_intensity_percent, 1) |
    plm::lag(sales_growth_1y_percent, 2:3) + plm::lag(rd_intensity_percent, 2:3),
  data = pdata,
  effect = "individual",
  model = "twosteps",
  transformation = "ld"
)
plm:::summary.pgmm(gmm_model)
```

Using a  System GMM model on a panel of 417 firms from 2010 to 2019, I found it examines the effect of R&D intensity on firm-level sales growth, while addressing endogeneity and dynamic panel bias. The coefficient on lagged R&D intensity  is estimated at 0.000138 and is highly statistically significant (z = 632.1, p < 0.001). This implies that a 1 percentage point increase in R&D intensity is associated with a 0.0138% increase in one-year sales growth, all else equal. Though the magnitude appears small, this is consistent with the scale and units of R&D intensity and the specification in levels.

Lagged firm growth  enters with a negative and highly significant coefficient (–0.00268), indicating mean reversion in growth. Meanwhile, current and lagged firm size are both significant, with signs suggesting that changes in size are positively related to future growth.

Model diagnostics support the reliability of the GMM estimation:
The Sargan test yields a p-value < 0.001, indicating possible instrument proliferation, which is a known risk in GMM; however,
The AR(2) test (p = 0.93) suggests no second-order autocorrelation, satisfying a critical assumption for GMM validity.
The Wald test strongly rejects the null that coefficients are jointly zero (p < 2.2e-16).

Overall, the results confirm that R&D intensity has a positive and statistically robust causal effect on firm growth, even when controlling for firm dynamics and unobserved heterogeneity. The GMM framework strengthens this interpretation by mitigating simultaneity and measurement error concerns that may bias OLS estimates.

2.R&D growth
```{r}
gmm_model_growth <- plm::pgmm(
  sales_growth_1y_percent ~ 
    plm::lag(sales_growth_1y_percent, 1) + 
    plm::lag(sales_million, 0:1) + 
    plm::lag(rd_growth_1y_percent, 1) |
    
    plm::lag(sales_growth_1y_percent, 2:3) + 
    plm::lag(rd_growth_1y_percent, 2:3),
    
  data = pdata,
  effect = "individual",
  model = "twosteps",
  transformation = "ld"
)

plm:::summary.pgmm(gmm_model_growth)
```

3.R&D log level
```{r}
merged_data <- merged_data %>%
  mutate(
    rd_log = log1p(rd_million),
    company = as.factor(company),
    year = as.integer(year)
  )

pdata <- pdata.frame(merged_data, index = c("company", "year")) %>%
  na.omit()

gmm_model_log <- plm::pgmm(
  sales_growth_1y_percent ~ 
    plm::lag(sales_growth_1y_percent, 1) + 
    plm::lag(sales_million, 0:1) + 
    plm::lag(rd_log, 1) |
    
    plm::lag(sales_growth_1y_percent, 2:3) + 
    plm::lag(rd_log, 2:3),
    
  data = pdata,
  effect = "individual",
  model = "twosteps",
  transformation = "ld"
)
plm:::summary.pgmm(gmm_model_log)
```

## LP-OLS

1.R&D intensity
```{r}
merged_data1 <- merged_data %>%
  mutate(
    company = as.factor(company),
    year = as.factor(year),
    country = as.factor(country),
    icb_sector = as.factor(icb_sector)
  )

contemp_data <- model.matrix(~ sales_million + year + country + icb_sector, data = merged_data1)[, -1]
contemp_data <- as.data.frame(contemp_data)


endog_data <- data.frame(
  sales_growth_1y_percent = merged_data$sales_growth_1y_percent,
  rd_intensity_percent = merged_data$rd_intensity_percent
)

results_lp <- lp_lin(
  endog_data    = endog_data,
  lags_endog_lin = 0,
  shock_type     = 1,
  trend          = 0,
  confint        = 1.96,
  hor            = 5,
  contemp_data   = contemp_data  
)

summary(results_lp)
plot(results_lp)
```

TUsing the local projection method (Jordà, 2005), we estimate the dynamic effects between R&D intensity and sales growth over a 5-year horizon. The results show that a one-percentage-point shock to R&D intensity leads to a persistent and statistically significant increase in firm growth from year 1 to year 5, with p-values < 0.001 and the largest F-statistic reaching 2.39 at horizon 5. The effect peaks around year 3–5, indicating that innovation efforts have cumulative returns over time. 

In contrast, sales growth does not significantly predict future R&D intensity across all horizons: all corresponding p-values are approximately 1.0, and the impulse response functions are flat and centered around zero. This asymmetry suggests a directional relationship—R&D drives growth, but not the other way around—helping to address concerns about reverse causality.

2.R&D growth
```{r}
endog_data_growth <- data.frame(
  sales_growth_1y_percent = merged_data$sales_growth_1y_percent,
  rd_growth_1y_percent = merged_data$rd_growth_1y_percent
)

results_lp_growth <- lp_lin(
  endog_data     = endog_data_growth,
  lags_endog_lin = 0,
  shock_type     = 1,
  trend          = 0,
  confint        = 1.96,
  hor            = 5,
  contemp_data   = contemp_data
)

summary(results_lp_growth)
plot(results_lp_growth)

```

3.R&D log level
```{r}
endog_data_log <- data.frame(
  sales_growth_1y_percent = merged_data$sales_growth_1y_percent,
  rd_log = merged_data$rd_log
)

results_lp_log <- lp_lin(
  endog_data     = endog_data_log,
  lags_endog_lin = 0,
  shock_type     = 1,
  trend          = 0,
  confint        = 1.96,
  hor            = 5,
  contemp_data   = contemp_data
)

summary(results_lp_log)
plot(results_lp_log)
```

## LP-2SLS

```{r}
#endog_data <- data.frame(
#  sales_growth_1y_percent = merged_data1$sales_growth_1y_percent,
# rd_intensity_percent = merged_data1$rd_intensity_percent)
# shock_var <- endog_data$rd_intensity_percent

# instrum_var <- data.frame(patent_intensity_sales = as.numeric(merged_data1$patent_intensity_sales))
# instrum_var <- data.frame(patent = as.numeric(merged_data1$patent_intensity_sales + 0))

# contemp_data <- model.matrix(~ sales_million + year + country + icb_sector, data = merged_data1)[, -1]
# contemp_data <- as.data.frame(contemp_data)

# results_lp_iv <- lp_lin_iv(
#  endog_data      = endog_data,
#  lags_endog_lin  = 0,
#  shock           = shock_var,
#  instrum         = instrum_var,
#  contemp_data    = contemp_data,
#  use_twosls      = TRUE,
#  trend           = 0,
#  confint         = 1.96,
#  hor             = 5)

# summary(results_lp_iv)
# plot(results_lp_iv)
```

I tried to run the two-stage local projection model using lp_lin_iv() from the lpirfs package, following the official documentation and examples closely. While the baseline model (lp_lin) runs fine, the 2SLS version repeatedly gives the error:

“Tool must be given in data.frame format”.

Even when I pass in the tool (patent_intensity_sales) as a properly formatted data frame (no NA, correct length, confirmed with str() and class()), the function still fails. I also tested a minimal example and reinstalled the latest GitHub version of lpirfs, but the problem persists.  /(TOT)/~~ ORZ

## Summary Table

1.Three indicators of OLS
```{r}
etable(ols_model, ols_model_growth, ols_model_log,
       dict = c(rd_lag1 = "R&D (lagged)",
                growth_lag1 = "Lagged Growth",
                size_lag1 = "Lagged Size"),
       headers = c("Intensity", "Growth", "Log Level"),
       title = "Comparison of R&D Measures in OLS Regressions",
       tex = FALSE)  
```

2.Three indicators of OLS-IV
```{r}
etable(iv_model, iv_model_growth, iv_model_log,
       dict = c(rd_lag1 = "R&D (Instrumented)",
                growth_lag1 = "Lagged Growth",
                size_lag1 = "Lagged Size"),
       headers = c("Intensity", "Growth", "Log Level"),
       title = "Comparison of IV Estimates Using Alternative R&D Measures",
       tex = FALSE) 
```

3.OLS,2SLS and GMM
```{r}
ols_tbl <- tidy(ols_model) |> mutate(Model = "OLS")
iv_tbl  <- tidy(iv_model)  |> mutate(Model = "IV (2SLS)")
gmm_sum <- summary(gmm_model)
gmm_tbl <- as.data.frame(gmm_sum$coefficients)
gmm_tbl <- gmm_tbl %>%
  tibble::rownames_to_column("term") %>%
  rename(estimate = Estimate, std.error = `Std. Error`, p.value = `Pr(>|z|)`) %>%
  mutate(Model = "GMM") %>%
  select(term, estimate, std.error, p.value, Model)
all_models <- bind_rows(ols_tbl, iv_tbl, gmm_tbl)
final_table <- all_models %>%
  filter(term %in% c("rd_lag1", "growth_lag1", "size_lag1", 
                     "plm::lag(rd_intensity_percent, 1)", 
                     "plm::lag(sales_growth_1y_percent, 1)", 
                     "plm::lag(sales_million, 0:1)0", 
                     "plm::lag(sales_million, 0:1)1")) %>%
  mutate(
    Variable = case_when(
      term %in% c("rd_lag1", "plm::lag(rd_intensity_percent, 1)") ~ "R&D (lagged)",
      term %in% c("growth_lag1", "plm::lag(sales_growth_1y_percent, 1)") ~ "Lagged Growth",
      term %in% c("size_lag1", "plm::lag(sales_million, 0:1)0") ~ "Lagged Size (t)",
      term == "plm::lag(sales_million, 0:1)1" ~ "Lagged Size (t-1)"
    ),
    Signif = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE            ~ ""
    ),
    Coef_SE = paste0(sprintf("%.6f", estimate), " (", sprintf("%.6f", std.error), ")", Signif)
  ) %>%
  select(Model, Variable, Coef_SE) %>%
  tidyr::pivot_wider(names_from = Model, values_from = Coef_SE) %>%
  relocate(Variable)
final_table %>%
  kbl(
    caption = "Comparison of OLS, IV, and GMM Estimates",
    align = "lccc"
  ) %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover")) %>%
  column_spec(1, bold = TRUE)

```

